---
title: "individual1"
author: "Alicia Hauglie"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("HAPPY.rds")
```

finrela <- opinion of family income
polviews <- think of self as liberal or conservative
wttsall <- weight variable
Inapplicable (IAP) <- respondents who are not asked to answer a specific question 

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
library(dplyr)
library(ggplot2)
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
HAPPY[HAPPY == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
HAPPY <- HAPPY %>% mutate_if(is.character, as.factor) %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89")) %>% mutate(AGE = as.numeric(levels(AGE)[AGE]))
str(HAPPY$AGE)
```
3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}
HAPPY <- HAPPY %>% mutate(MARITAL = factor(MARITAL, levels = c("NEVER MARRIED", "MARRIED", "SEPARATED", "DIVORCED", "WIDOWED"))) %>% arrange(desc(MARITAL))

HAPPY <- HAPPY %>% mutate(HAPPY = factor(HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))) %>% arrange(desc(HAPPY))

HAPPY <- HAPPY %>% mutate(DEGREE = factor(DEGREE, levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))) %>% arrange(desc(DEGREE))

HAPPY <- HAPPY %>% mutate(FINRELA = factor(FINRELA, levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))) %>% arrange(desc(FINRELA))

HAPPY <- HAPPY %>% mutate(HEALTH = factor(HEALTH, levels = c("POOR", "FAIR", "GOOD", "EXCELLENT"))) %>% arrange(desc(HEALTH))

HAPPY <- HAPPY %>% mutate(PARTYID = factor(PARTYID, levels = c("STRONG DEMOCRAT", "NOT STR DEMOCRAT", "IND,NEAR DEM", "INDEPENDENT", "IND,NEAR REP", "NOT STR REPUBLICAN", "STRONG REPUBLICAN", "OTHER PARTY"))) %>% arrange(desc(PARTYID))

HAPPY <- HAPPY %>% mutate(POLVIEWS = factor(POLVIEWS, levels = c("EXTREMELY LIBERAL", "LIBERAL", "SLIGHTLY LIBERAL", "MODERATE", "SLIGHTLY CONSERVATIVE", "CONSERVATIVE", "EXTRMLY CONSERVATIVE"))) %>% arrange(desc(POLVIEWS))

```

**PART TWO**
1. Investigate the distributions of 2 different variables.
```{r}
ggplot(data=HAPPY, aes(x=DEGREE))+geom_bar(fill="darkgreen")+facet_wrap(~SEX)+coord_flip()
```
this shows us the distribution of education between the sexes.
```{r}
ggplot(data=HAPPY, aes(x=AGE))+geom_histogram(binwidth=1, fill="darkorange")+facet_wrap(~SEX)

```
this shows us the age distribution between the sexes, it also shows that more women took this survey than men. 



2. Investigate the relationship between happiness and three other variables in the data.
```{r}
library(ggmosaic)
ggplot(data=HAPPY)+geom_mosaic(aes(x=product(AGE),fill=SEX))+facet_wrap(HAPPY$HAPPY)


ggplot(data=HAPPY)+geom_boxplot(aes=FINRELA)
ggplot(data=HAPPY, aes(y=(AGE)))+geom_boxplot()+facet_wrap(HAPPY$FINRELA)+coord_flip()

library(ggmosaic)
ggplot(data=HAPPY)+geom_mosaic(aes(x=product(HEALTH),fill=FINRELA))

```
3. What do you learn from the data set? Write one or two paragraphs to describe.
```{r}

```
Explore the dataset yourself, and then discuss and refine answers as a team.

