---
title: "individual2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("HAPPY.rds")
library(plyr)
library(ggmosaic)
library(dplyr)
library(ggthemes)
```

finrela <- Opinion of family income
polviews <- Think of self as liberal or conservative
wttsall <- weight variable
Inapplicable (IAP). Respondents who are not asked to answer a specific question are assigned to IAP.

Answer each of the following questions using R code to get an answer. Write a separate code chunk to manipulate each variable. Once all the data manipulation is done, show the values in a barchart or histogram and comment on the plot. Use ggplot2 to create the plots.

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
library(dplyr)
library(ggplot2)
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
HAPPY[HAPPY == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
HAPPY <- HAPPY %>% mutate_if(is.character, as.factor) %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89")) %>% mutate(AGE = as.numeric(levels(AGE)[AGE]))
str(HAPPY$AGE)
str(HAPPY)
```

3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}
HAPPY <- HAPPY %>% mutate(MARITAL = factor(MARITAL, levels = c("NEVER MARRIED", "MARRIED", "SEPARATED", "DIVORCED", "WIDOWED"))) %>% arrange(desc(MARITAL))

 HAPPY <- HAPPY %>% mutate( HAPPY= factor(HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))) %>% arrange(desc(HAPPY))
 
HAPPY <- HAPPY %>%
  mutate( DEGREE= factor(DEGREE, 
                          levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))) %>% arrange(desc(DEGREE))

HAPPY <- HAPPY %>%
  mutate( FINRELA= factor(FINRELA, 
                          levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))) %>% 
  arrange(desc(FINRELA))

HAPPY <- HAPPY %>%
  mutate( HEALTH= factor(HEALTH, 
                          levels = c("POOR ", "FAIR", "GOOD", "EXCELLENT"))) %>% arrange(desc(HEALTH))

HAPPY <- HAPPY %>%
  mutate( PARTYID= factor(PARTYID, 
                          levels = c("STRONG DEMOCRAT", "NOT STR DEMOCRAT", "IND,NEAR DEM", "INDEPENDENT", "IND,NEAR REP", "NOT STR REPUBLICAN","STRONG REPUBLICAN", "OTHER PARTY"))) %>% arrange(desc(PARTYID))

HAPPY <- HAPPY %>%
  mutate( POLVIEWS= factor(POLVIEWS, 
                          levels = c("EXTREMELY LIBERAL", "LIBERAL", "SLIGHTLY LIBERAL", "MODERATE", "SLIGHTLY CONSERVATIVE", "CONSERVATIVE","EXTRMELY CONSERVATIVE"))) %>% arrange(desc(POLVIEWS))
levels(HAPPY$POLVIEWS)
```

Part 2
Happiness!
Use ggplot2 to create the plots. Your exploration should make use all elements of the grammar of graphics: mappings (color, fill, and size at least), different geoms (4 at least), scales, coordinates (eg coord_flip), facet, and theme (eg setting axis labels and title). For each investigation, create an appropriate visualization and write a paragraph to describe it.

1. Investigate the distributions of 2 different variables.
```{r}
ggplot(data=HAPPY,aes(x=MARITAL))+ geom_bar(fill="lightblue") + facet_wrap(~SEX) + coord_flip()

ggplot(HAPPY, aes(x= HAPPY)) + geom_bar( aes(fill=..count..)) + scale_fill_gradient("Count", low="green", high ="red") 

ggplot(HAPPY, aes(x=AGE, y=HAPPY)) + geom_count()


pie <- ggplot(HAPPY, aes(x = "", fill = factor(DEGREE ))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of Degree", 
       caption="Source: HAPPY")
  
pie + coord_polar(theta = "y", start=0)

#This pie chart is weird, it shows that more than half of the people in this data set have a highschool degree, and the average age is above 25, meaning most people in this study didnt go to college, and probably arent planning on it. 
```

2. Investigate the relationship between happiness and three other variables in the data.
```{r}
ggplot(HAPPY, aes(x=SEX)) +geom_bar() + facet_wrap(~HAPPY)
ggplot(HAPPY, aes(x=HEALTH)) +geom_bar() + facet_wrap(~HAPPY)
ggplot(HAPPY, aes(x=DEGREE)) +geom_bar() + facet_wrap(~HAPPY) + coord_flip()

ggplot(HAPPY, aes(x=AGE, fill=SEX)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY)
# THE COUNT OF HAPPINESS SEEMS VERY EVEN THROUGHOUT THE SEXES, and shows that happiness doesn't depend on age as each graph has the same shape.



ggplot(HAPPY, aes(x=AGE, fill=FINRELA)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY) + ggtitle("Plot of Happiness with Age and Finrela (Opinion of family income)") +
  xlab("Age") + ylab("Count")
# This is a Hitogram of Age which is facet wrapped with Happiness. To see the breakdown of how money affects happiness I decided to break down the fill by peoples own opinion of there income. This is the first plot that I have made that i can get some conlclusions about what increases happiness. As you can see the highest two categories of Finrela, Above Average and Far Above Average, are barely in the "NOT TOO HAPPY" plot. In contrast, the top two colums in Finrela, show a lot larger portion of the "PRETTY HAPPY" and "VERY HAPPY" plots, meaning that if peoples opinion of there income is high, there happiness is likely to be higher. 


ggplot(HAPPY, aes(x=YEAR, fill=DEGREE)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY) + coord_flip()


p <- ggplot(HAPPY, aes(x=HAPPY, y=AGE)) + geom_boxplot()
p + ggtitle("Plot of happiness by Age") +
  xlab("Happiness") + ylab("age")

ggplot(HAPPY, aes(x=HEALTH, y=AGE)) + geom_boxplot() +facet_wrap(~HAPPY)
# This plot is a numbber of boxplots that are broken down by age and health, then facet wrapped by health. As you can see the the plots show that as average age goes down, health goes up, which is logical in itself. We can also conclude that health does not have anything to do with happiness, as there is not an outlier of happiness with each level of health. Each level of happiness has the same shape of boxplots. 
```

3. What do you learn from the data set? Write one or two paragraphs to describe.
Explore the dataset yourself, and then discuss and refine answers as a team.

My conlcusions are that I learned that working with categorical data can be very frustring especially when we have really only one colummn of continous data, age. This was tough to find any correlations between happiness and other variables. I did, however, find that with the opinion of family income does have an effect on happiness, which makes sense logically. More money would probably result in more happiness. When looking at health, i didn't find a significance between health and happiness. 

Overall, I did not like using this data set. It seemed very skewed that 3/4 of the data was people who had only a highschool degree or lower. This might have been closer to the actual population in the 1970's when this was taken, but as of the last few years this seems very low to represent the population.








