---
title: "individual2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("HAPPY.rds")
library(plyr)
install.packages('ggthemes', dependencies = TRUE)
library(ggmosaic)
library(dplyr)
library(ggthemes)
```

finrela <- Opinion of family income
polviews <- Think of self as liberal or conservative
wttsall <- weight variable
Inapplicable (IAP). Respondents who are not asked to answer a specific question are assigned to IAP.

Answer each of the following questions using R code to get an answer. Write a separate code chunk to manipulate each variable. Once all the data manipulation is done, show the values in a barchart or histogram and comment on the plot. Use ggplot2 to create the plots.

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
library(dplyr)
library(ggplot2)
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
HAPPY[HAPPY == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
HAPPY <- HAPPY %>% mutate_if(is.character, as.factor) %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89")) %>% mutate(AGE = as.numeric(levels(AGE)[AGE]))
str(HAPPY$AGE)
str(HAPPY)
```

3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}
HAPPY <- HAPPY %>% mutate(MARITAL = factor(MARITAL, levels = c("NEVER MARRIED", "MARRIED", "SEPARATED", "DIVORCED", "WIDOWED"))) %>% arrange(desc(MARITAL))

 HAPPY <- HAPPY %>% mutate( HAPPY= factor(HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))) %>% arrange(desc(HAPPY))
 
HAPPY <- HAPPY %>%
  mutate( DEGREE= factor(DEGREE, 
                          levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))) %>% arrange(desc(DEGREE))

HAPPY <- HAPPY %>%
  mutate( FINRELA= factor(FINRELA, 
                          levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))) %>% 
  arrange(desc(FINRELA))

HAPPY <- HAPPY %>%
  mutate( HEALTH= factor(HEALTH, 
                          levels = c("POOR ", "FAIR", "GOOD", "EXCELLENT"))) %>% arrange(desc(HEALTH))

HAPPY <- HAPPY %>%
  mutate( PARTYID= factor(PARTYID, 
                          levels = c("STRONG DEMOCRAT", "NOT STR DEMOCRAT", "IND,NEAR DEM", "INDEPENDENT", "IND,NEAR REP", "NOT STR REPUBLICAN","STRONG REPUBLICAN", "OTHER PARTY"))) %>% arrange(desc(PARTYID))

HAPPY <- HAPPY %>%
  mutate( POLVIEWS= factor(POLVIEWS, 
                          levels = c("EXTREMELY LIBERAL", "LIBERAL", "SLIGHTLY LIBERAL", "MODERATE", "SLIGHTLY CONSERVATIVE", "CONSERVATIVE","EXTRMELY CONSERVATIVE"))) %>% arrange(desc(POLVIEWS))
levels(HAPPY$POLVIEWS)
```

Part 2
Happiness!
Use ggplot2 to create the plots. Your exploration should make use all elements of the grammar of graphics: mappings (color, fill, and size at least), different geoms (4 at least), scales, coordinates (eg coord_flip), facet, and theme (eg setting axis labels and title). For each investigation, create an appropriate visualization and write a paragraph to describe it.

1. Investigate the distributions of 2 different variables.
```{r}
ggplot(data=HAPPY, aes(x = AGE, main = "Histogram for Age" )) + geom_histogram(binwidth = 5)
ggplot(data=HAPPY,aes(x=MARITAL))+ geom_bar(fill="lightblue") + facet_wrap(~SEX) + coord_flip()

ggplot(HAPPY, aes(x= HAPPY)) + geom_bar( aes(fill=..count..)) + scale_fill_gradient("Count", low="green", high ="red") 
ggplot(HAPPY, aes(x=AGE)) + geom_area(binwidth = 1) + facet_wrap(~YEAR)

ggplot(HAPPY, aes(x=AGE, y=HAPPY)) + geom_count()

ggplot(HAPPY, aes(x=AGE, y=HAPPY)) + geom()




```

2. Investigate the relationship between happiness and three other variables in the data.
```{r}
ggplot(HAPPY, aes(x=SEX)) +geom_bar() + facet_wrap(~HAPPY)
ggplot(HAPPY, aes(x=HEALTH)) +geom_bar() + facet_wrap(~HAPPY)
ggplot(HAPPY, aes(x=DEGREE)) +geom_bar() + facet_wrap(~HAPPY) + coord_flip()

ggplot(HAPPY, aes(x=AGE, fill=SEX)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY)
# THE COUNT OF HAPPINESS SEEMS VERY EVEN THROUGHOUT THE SEXES, and shows that happiness doesn't depend on age as each graph has the same shape.



ggplot(HAPPY, aes(x=AGE, fill=FINRELA)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY) + ggtitle("Plot of Happiness with Age and Finrela (Opinion of family income)") +
  xlab("Age") + ylab("Count")
# This is a Hitogram of Age which is facet wrapped with Happiness. To see the breakdown of how money affects happiness I decided to break down the fill by peoples own opinion of there income. This is the first plot that I have made that i can get some conlclusions about what increases happiness. As you can see the highest two categories of Finrela, Above Average and Far Above Average, are barely in the "NOT TOO HAPPY" plot. In contrast, the top two colums in Finrela, show a lot larger portion of the "PRETTY HAPPY" and "VERY HAPPY" plots, meaning that if peoples opinion of there income is high, there happiness is likely to be higher. 


ggplot(HAPPY, aes(x=YEAR, fill=DEGREE)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY) + coord_flip()

pie <- ggplot(HAPPY, aes(x = "", fill = factor(DEGREE ))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of Degree", 
       caption="Source: HAPPY")
  
pie + coord_polar(theta = "y", start=0)
```

3. What do you learn from the data set? Write one or two paragraphs to describe.
Explore the dataset yourself, and then discuss and refine answers as a team.
```{r}

```





