---
title: "team"
author: "Alicia Hauglie & Samuel Cole"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("HAPPY.rds")
```

finrela <- opinion of family income
polviews <- think of self as liberal or conservative
wttsall <- weight variable
Inapplicable (IAP) <- respondents who are not asked to answer a specific question 

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
library(dplyr)
library(ggplot2)
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
HAPPY[HAPPY == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
HAPPY <- HAPPY %>% mutate_if(is.character, as.factor) %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89")) %>% mutate(AGE = as.numeric(levels(AGE)[AGE]))
str(HAPPY$AGE)
```
3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.


 **PART TWO**
1. Investigate the distributions of 2 different variables.
```{r}

```
2. Investigate the relationship between happiness and three other variables in the data.
```{r}

```
3. What do you learn from the data set? Write one or two paragraphs to describe.
```{r}

```
Explore the dataset yourself, and then discuss and refine answers as a team.

