---
title: "team"
author: "Alicia Hauglie & Samuel Cole"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("HAPPY.rds")
```

finrela <- opinion of family income
polviews <- think of self as liberal or conservative
wttsall <- weight variable
Inapplicable (IAP) <- respondents who are not asked to answer a specific question 

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
library(dplyr)
library(ggplot2)
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
HAPPY[HAPPY == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
HAPPY <- HAPPY %>% mutate_if(is.character, as.factor) %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89")) %>% mutate(AGE = as.numeric(levels(AGE)[AGE]))
str(HAPPY$AGE)
```
3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.

```{r}
HAPPY <- HAPPY %>% mutate(MARITAL = factor(MARITAL, levels = c("NEVER MARRIED", "MARRIED", "SEPARATED", "DIVORCED", "WIDOWED"))) %>% arrange(desc(MARITAL))

 HAPPY <- HAPPY %>% mutate( HAPPY= factor(HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))) %>% arrange(desc(HAPPY))
 
HAPPY <- HAPPY %>%
  mutate( DEGREE= factor(DEGREE, 
                          levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))) %>% arrange(desc(DEGREE))

HAPPY <- HAPPY %>%
  mutate( FINRELA= factor(FINRELA, 
                          levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))) %>% 
  arrange(desc(FINRELA))

HAPPY <- HAPPY %>%
  mutate( HEALTH= factor(HEALTH, 
                          levels = c("POOR ", "FAIR", "GOOD", "EXCELLENT"))) %>% arrange(desc(HEALTH))

HAPPY <- HAPPY %>%
  mutate( PARTYID= factor(PARTYID, 
                          levels = c("STRONG DEMOCRAT", "NOT STR DEMOCRAT", "IND,NEAR DEM", "INDEPENDENT", "IND,NEAR REP", "NOT STR REPUBLICAN","STRONG REPUBLICAN", "OTHER PARTY"))) %>% arrange(desc(PARTYID))

HAPPY <- HAPPY %>%
  mutate( POLVIEWS= factor(POLVIEWS, 
                          levels = c("EXTREMELY LIBERAL", "LIBERAL", "SLIGHTLY LIBERAL", "MODERATE", "SLIGHTLY CONSERVATIVE", "CONSERVATIVE","EXTRMELY CONSERVATIVE"))) %>% arrange(desc(POLVIEWS))
levels(HAPPY$POLVIEWS)
```

 **PART TWO**
1. Investigate the distributions of 2 different variables.
```{r}
ggplot(HAPPY, aes(x= HAPPY)) + geom_bar( aes(fill=..count..)) + scale_fill_gradient("Count", low="green", high ="red") + ggtitle("Distribution of Happiness") +
  xlab("HAPPY") + ylab("Count")
```
This graph shows the overall distribution of happiness without accounting for any other variable. We can easily see, by the color, that PRETTY HAPPY, is the most selected choice, and VERY HAPPY is the second most selected choice. This makes sense to me, as people want to be happy, so the distribution would favor the happy side. So overall, we can see that around half of the participants chose PRETTY HAPPY as there level of happiness. 

```{r}
pie <- ggplot(HAPPY, aes(x = "", fill = factor(DEGREE ))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of Degree", 
       caption="Source: HAPPY")
  
pie + coord_polar(theta = "y", start=0)
```
This graph shows us the education distribution of the people who took the survey. An interesting point is that over half of people who took this survey only have a high school education, even though most of the people surveyed are over the age of 25. It might be good to point out that this survey was taken years ago because if you were to survey random people above the age of 25 nowadays, you would find a much larger proportion of people who have had at least some college education. This is probably not a very representative sample, as only about 30% of the survey takers have more than a high school education. 

2. Investigate the relationship between happiness and three other variables in the data.
```{r}
ggplot(HAPPY, aes(x=AGE, fill=FINRELA)) + geom_histogram(binwidth = 5) + facet_grid(~HAPPY, scales= 'free_y') + ggtitle("Plot of Happiness with Age and Finrela (Opinion of family income)") +
  xlab("Age") + ylab("Count")
```
This is a Histogram of Age which is facet wrapped with Happiness. To see the breakdown of how money affects happiness I decided to break down the fill by peoples own opinion of there income. This is the first plot that I have made that i can get some conlclusions about what increases happiness. As you can see the highest two categories of Finrela, Above Average and Far Above Average, are barely in the "NOT TOO HAPPY" plot. In contrast, the top two colums in Finrela, show a lot larger portion of the "PRETTY HAPPY" and "VERY HAPPY" plots, meaning that if peoples opinion of there income is high, there happiness is likely to be higher. 

```{r}
library(ggmosaic)
ggplot(data=HAPPY)+geom_mosaic(aes(x=product(HEALTH),fill=FINRELA))
```
This graph shows us how the survey takers financial situations are distributed in relation to how they said their health was. As you can see, no one who took the survey said they had poor health. The graph shows that people are more likely to answer that they are only in FAIR health when their financial situation falls in the FAR BELOW AVERAGE or BELOW AVERAGE category. We might conclude that people in worse financial situations are less happy, possibly because of the stress that comes with this. This also shows that less than 20% of the respondents said they were in FAIR health. People who answered EXCELLENT health are also more likely to have a finrela in the ABOVE AVERAGE or FAR ABOVE AVERAGE category. 

```{r}
ggplot(HAPPY, aes(x=HEALTH, y=AGE)) + geom_boxplot() +facet_wrap(~HAPPY)
```
This plot is a numbber of boxplots that are broken down by age and health, then facet wrapped by health. As you can see the the plots show that as average age goes down, health goes up, which is logical in itself. We can also conclude that health does not have anything to do with happiness, as there is not an outlier of happiness with each level of health. Each level of happiness has the same shape of boxplots.

3. What do you learn from the data set? Write one or two paragraphs to describe.

My conlcusions are that I learned that working with categorical data can be very frustring especially when we have really only one colummn of continous data, age. This was tough to find any correlations between happiness and other variables. I did, however, find that with the opinion of family income does have an effect on happiness, which makes sense logically. More money would probably result in more happiness. When looking at health, i didn't find a significance between health and happiness. 

Overall, I did not like using this data set. It seemed very skewed that 3/4 of the data was people who had only a highschool degree or lower. This might have been closer to the actual population in the 1970's when this was taken, but as of the last few years this seems very low to represent the population.

